### 2014-01-31 analysis/Outline.rmd
updated scaling of fine root vs. leaf mass, and leaf mass fraction of abvground biomass

if(!file.exists("../cache/allomdata.rds")){
  dat <- loadStudies(reprocess=FALSE)
  data <- fillDerivedVariables(dat$data)
  saveRDS(data, file="cache/allomdata.rds")
} else {
  data <- readRDS("../cache/allomdata.rds")
}

@@

palette(rich.colors(4))
with(dfr, plot(log10(m.lf), log10(m.rf),
               axes=FALSE,
               xlab="Leaf biomass (kg)",
               ylab="Fine root biomass (kg)",
               pch=19,
               col=pft))
zs <- c(0.05, 0.25, 0.5, 0.75,1)
for(z in zs)abline(log10(z),1)
legend("bottomright", levels(dfr$pft), pch=19, col=palette())
magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
box()
```
Figure X. Lines are with exponent of 1, and represent fixed ratios between fine root biomass and leaf biomass (from top to bottom: 1, 0.75, 0.5, 0.25, 0.05).

@@

```{r echo=FALSE}
# subset/refine data.
dfr <- droplevels(subset(data, !is.na(m.st) & !is.na(m.lf) & pft != "" & m.lf > 0))
dfr <- within(dfr, {
  dataset <- as.factor(dataset)
  species <- as.factor(species)
  pft <- as.factor(pft)
  m.lf.fraction <- with(dfr, m.lf / m.so)
  log_m.so <- log10(m.so)
  log_m.lf.fraction <- log10(m.lf.fraction)
})

with(dfr, plot(log_m.so, log_m.lf.fraction, axes=FALSE,
               xlab="Aboveground biomass (kg)",
               ylab="Leaf mass fraction (-)",
               pch=19, col=pft
))
magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
box()
legend("topright", levels(dfr$pft), pch=19, col=palette())
```
Figure X. Much higher leaf mass fraction for evergreen gymnosperms at a given total aboveground plant biomass.

```{r echo=FALSE}
dfr <- droplevels(subset(data, !is.na(m.st) & !is.na(m.lf) & !is.na(m.rt) & 
                           m.lf > 0 & m.rt > 0 & pft != "" & 
                           m.st < m.to))
dfr$m.to <- with(dfr, m.rt + m.st + m.lf)

@@

>>>>

if(!file.exists("../cache/allomdata.rds")){
  dat <- loadStudies(reprocess=FALSE)
  data <- fillDerivedVariables(dat$data)
  saveRDS(data, file="cache/allomdata.rds")
} else {
  data <- readRDS("../cache/allomdata.rds")
}
data$dataset_species <- as.factor(paste(data$dataset, data$species, sep="-"))

@@

sm <- lapply(split(dfr, dfr$pft), function(x){
      sma(log10(m.rf) ~ log10(m.lf)*dataset_species, 
           data=x)
      })

palette(alpha(rich.colors(4),0.5))
with(dfr, plot(log10(m.lf), log10(m.rf), 
               axes=FALSE,
               xlab="Aboveground biomass (kg)",
               ylab="Foliage mass (kg)",
               type='n',col=pft, panel.first={
                 for(z in c(0.05, 0.25, 0.5, 0.75, 1))abline(log10(z),1)
               }))
magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
for(i in 1:3)plot(sm[[i]], add=TRUE, col=palette()[i], type='l', lwd=2)
box()
legend("bottomright", levels(dfr$pft), fill=alpha(palette(),1))

```
Figure X. Lines are with exponent of 1, and represent fixed ratios between fine root biomass and leaf biomass (from top to bottom: 1, 0.75, 0.5, 0.25, 0.05).

@@

```{r echo=FALSE, eval=FALSE}
# subset/refine data.
dfr <- droplevels(subset(data, !is.na(m.st) & !is.na(m.lf) & pft != "" & m.lf > 0))
dfr <- within(dfr, {
  dataset <- as.factor(dataset)
  species <- as.factor(species)
  pft <- as.factor(pft)
  m.lf.fraction <- with(dfr, m.lf / m.so)
  log_m.so <- log10(m.so)
  log_m.lf.fraction <- log10(m.lf.fraction)
})

with(dfr, plot(log_m.so, log_m.lf.fraction, axes=FALSE,
               xlab="Aboveground biomass (kg)",
               ylab="Leaf mass fraction (-)",
               pch=19, col=pft
))
magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
box()
legend("topright", levels(dfr$pft), pch=19, col=palette())
#Figure X. Much higher leaf mass fraction for evergreen gymnosperms at a given total aboveground plant biomass.
```

```{r echo=FALSE, eval=FALSE}
dfr <- droplevels(subset(data, !is.na(m.st) & !is.na(m.lf) & !is.na(m.rt) & 
                           m.lf > 0 & m.rt > 0 & pft != "" & 
                           m.st < m.to))
dfr$m.to <- with(dfr, m.rt + m.st + m.lf)

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}

dfr <- droplevels(subset(data, !is.na(m.so) & !is.na(m.lf) & pft != "" &
                           m.lf > 0 & m.so > 0))
dfr$pft <- as.factor(dfr$pft)

sm <- lapply(split(dfr, dfr$pft), function(x){
      sma(log10(m.lf) ~ log10(m.so)*dataset_species, 
           data=x)
      })

palette(alpha(rich.colors(4), 0.4))
with(dfr, plot(log10(m.so), log10(m.lf), 
               axes=FALSE,
               xlab="Aboveground biomass (kg)",
               ylab="Foliage mass (kg)",
               type='n',col=pft, panel.first={
                 for(z in c(-2,-1,0))abline(z,1)

               }))
magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
for(i in 1:4)plot(sm[[i]], add=TRUE, col=palette()[i], type='l')
box()
legend("bottomright", levels(dfr$pft), fill=alpha(palette(),1))
